---
title: "ABM Group Dynamics in CPS"
author: "Kristiane Uhrenholt Warncke and Anne Skamris Holm"
date: "2024-05-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
pacman::p_load(tidyverse, ggpubr, ggalt, ggExtra,  plotrix, dplyr, ggplot2, stringr, tidyr, gridExtra, grid, purrr, lme4, corrplot, patchwork, rstanarm, cowplot)
```

## Loading the data 
```{r}
df <- read.csv2("ABM_FINAL_FINAL.csv")
```

### PREPROCESSING

```{r}
#Renaming columns 
df <- rename(df, 
                     round = X.run.number., 
                     team_size = team_size, 
                     game_size = game_size,
                     know_sharing = knowledge_sharing,
                     n_steps = X.step.,
                     team_score = team_score,
                     gain_score = gain_score,
                     synergy_score = synergy_score,
                     agent_names = agent_names,
                     reached_consensus = reached_consensus.,
                     agents_agree = X.agreeableness..of.turtles,
                     agents_talk = X.talkativeness..of.turtles,
                     agents_neuro = X.neuroticism..of.turtles,
                     agents_crit = X.critical_thinking..of.turtles,
                     agents_know= X.knowledge..of.turtles)
```

## Converting string vectors to numeric vectors in certain columns
```{r}
# Helper function to convert string vectors to numeric vectors
convert_to_numeric_vector <- function(x) {
  # Remove unwanted characters
  cleaned <- gsub("\\[|\\]|\"", "", x)
  # Split the string by spaces and convert to numeric
  as.numeric(unlist(strsplit(cleaned, " ")))
}

# Apply this conversion to each agents_ column
df <- df %>%
  mutate(across(starts_with("agents_"), ~map(.x,convert_to_numeric_vector)))

# ..and to the synergy score
df <- df %>%
  mutate(across("synergy_score", ~map(.x, convert_to_numeric_vector)))

# and team score
df <- df %>%
  mutate(across("team_score", ~map(.x, convert_to_numeric_vector)))

# and gain score
df <- df %>%
  mutate(across("gain_score", ~map(.x, convert_to_numeric_vector)))

```
## Grouping the diversity_score into levels from 1-5 depending on team size 
```{r}
# Calculate diversity level
df <- df %>%
  mutate(
    agent_names_cleaned = str_extract_all(agent_names, "\\b\\w+_agent\\b"), # Extract agent types
    diversity_level = map_int(agent_names_cleaned, ~length(unique(.x))) # Count unique agent types
  )%>%
  select(-know_sharing, -team_size, -game_size, -agent_names)

#df$agent_names_cleaned <- lapply(df$agent_names_cleaned, sort)

# Function to remove "_agent" from each element in the list
clean_agent_names <- function(list) {
  sapply(list, function(x) gsub("_agent", "", x), USE.NAMES = FALSE)
}

# Apply the function to each list in the column and sort the cleaned lists
df$agent_names_cleaned <- lapply(df$agent_names_cleaned, function(x) sort(clean_agent_names(x)))
```

```{r}
df <- df %>%
  mutate(
    round = as.integer(round),
    n_steps = as.numeric(n_steps),
    team_score = as.numeric(team_score),
    gain_score = as.numeric(gain_score),
    synergy_score = as.numeric(synergy_score),
    agent_names_cleaned = as.character(agent_names_cleaned),
    reached_consensus = as.logical(reached_consensus),  # this is a TRUE/FALSE column
    diversity_level = as.factor(diversity_level)
  )

df <- df %>%
  drop_na() %>%
  mutate(agree_dom = as.numeric(lapply(agents_agree, mean)))%>%
  mutate(talk_dom = as.numeric(lapply(agents_talk, mean)))%>%
  mutate(crit_dom = as.numeric(lapply(agents_crit, mean)))%>%
  mutate(know_dom = as.numeric(lapply(agents_know, mean)))%>%
  mutate(neuro_dom = as.numeric(lapply(agents_neuro, mean)))

df <- df %>%
  mutate(agree_dom_0=agree_dom)%>%
  mutate(talk_dom_0 =talk_dom)%>%
  mutate(crit_dom_0 =crit_dom)%>%
  mutate(know_dom_0 =know_dom)%>%
  mutate(neuro_dom_0 =neuro_dom)

df <- df %>%
  mutate(agree_dom= ifelse(agree_dom==0, NA, agree_dom))%>%
  mutate(talk_dom = ifelse(talk_dom==0, NA, talk_dom))%>%
  mutate(crit_dom = ifelse(crit_dom==0, NA, crit_dom))%>%
  mutate(know_dom = ifelse(know_dom==0, NA, know_dom))%>%
  mutate(neuro_dom = ifelse(neuro_dom==0, NA, neuro_dom))

df <- df %>%
  mutate(agree_dom_std = scale(df$agree_dom))%>%
  mutate(talk_dom_std = scale(talk_dom))%>%
  mutate(crit_dom_std = scale(crit_dom))%>%
  mutate(know_dom_std = scale(know_dom))%>%
  mutate(neuro_dom_std = scale(neuro_dom))

df <- df %>%
  select(round, n_steps, team_score, gain_score, synergy_score, reached_consensus, diversity_level, agents_agree,agree_dom_0, agree_dom, agree_dom_std,  agents_talk, talk_dom_0,talk_dom,talk_dom_std,  agents_crit,crit_dom_0, crit_dom,crit_dom_std,agents_know,know_dom_0,know_dom,know_dom_std,agents_neuro, neuro_dom_0, neuro_dom,neuro_dom_std, agent_names_cleaned)%>%
  filter(reached_consensus=="TRUE")
```


## Inspecting Data
```{r}
#Team, gain and synergy plots
synergyscore_plot <- ggplot(df, aes(x = synergy_score)) +
  geom_histogram(bins = 30, fill = "skyblue", color = "black") +
  theme_minimal() +
  labs(title = "Distribution of Group Performance on the three score-types", x = "Synergy Score", y = "Frequency")

teamscore_plot <- ggplot(df, aes(x = team_score)) +
  geom_histogram(bins = 30, fill = "lightgreen", color = "black") +
  theme_minimal() +
  labs(title = "", x = "Team Score", y = "")

gainscore_plot <- ggplot(df, aes(x = gain_score)) +
  geom_histogram(bins = 30, fill = "lightpink", color = "black") +
  theme_minimal() +
  labs(title = "", x = "Gain score", y = "")

grid.arrange(synergyscore_plot, teamscore_plot, gainscore_plot, ncol=3)
```

```{r}
df%>% 
  group_by(diversity_level) %>% 
  summarise("count" = n())
```


```{r}
#distribution of agents
colSums(!is.na(df))[8:22]
```

### density plots....

```{r}
#long data to make density plots across traits
long_data <- df %>%
  pivot_longer(
    cols = c("talk_dom_std", "neuro_dom_std", "crit_dom_std", "know_dom_std", "agree_dom_std"),
    names_to = "trait",
    values_to = "value"
  ) %>%
  drop_na(value)

density_plot_combined <- ggplot(long_data, aes(x = value, fill = trait, color = trait)) +
  geom_density(alpha = 0.7) +  # Adding transparency to see overlapping areas
  scale_fill_brewer(palette = "Paired") +  # This sets distinct colors for each trait
  scale_color_brewer(palette = "Paired") +
  theme_minimal() +
  labs(
    title = "Distribution of Scaled Dominating Traits",
    x = "Trait Value",
    y = "Density"
  ) +
  scale_x_continuous(limits = c(-4, 6)) +
  theme(legend.title = element_blank())  # Optionally remove the legend title

density_plot_combined
```

```{r}
ggplot(long_data, aes(x = trait, y = value, fill = trait)) +
  geom_boxplot() +
  scale_fill_brewer(palette = "Set3") +  # Using a color palette for distinction
  theme_minimal() +
  labs(title = "Boxplots of Trait Values", x = "Trait", y = "Value") +
  scale_y_continuous(limits = c(-2, 7)) +
  theme(legend.position = "none")  # Optional: remove the legend if not needed
```
```{r}
#Calculating average synergy, team and gain scores
average_synergy_score <- mean(df$synergy_score)
average_team_score <- mean(df$team_score)
average_gain_score <- mean(df$gain_score)

# Print the result
print(paste("Average synergy score:", average_synergy_score))
print(paste("Average team score:", average_team_score))
print(paste("Average gain score:", average_gain_score))

```

```{r}
mean_agree <- mean(df$agree_dom, na.rm = TRUE)
mean_talk <- mean(df$talk_dom, na.rm = TRUE)
mean_neuro <- mean(df$neuro_dom, na.rm = TRUE)
mean_crit <- mean(df$crit_dom, na.rm = TRUE)
mean_know <- mean(df$know_dom, na.rm = TRUE)

# Print the results
print(paste("Average agree:", mean_agree))
print(paste("Average talk:", mean_talk))
print(paste("Average neuro:", mean_neuro))
print(paste("Average crit:", mean_crit))
print(paste("Average know:", mean_know))
```

```{r}
# Convert each list to a character string
df$diversity_level<-as.factor(df$diversity_level)

df$lists_str <- sapply(df$agent_names_cleaned, function(x) paste(sort(x), collapse = ","))

# Create a frequency table of unique lists by diversity_level
freq_table <- df %>%
  group_by(lists_str, diversity_level)%>%
  summarise(gain_score=mean(gain_score))

freq_table2 <- df %>%
  group_by(diversity_level, lists_str) %>%
  summarise(Frequency = n(), .groups = 'drop')

# Create the grouped bar plot using ggplot2 without labels
p2 <- ggplot(freq_table2, aes(x = reorder(lists_str, -Frequency), y = Frequency, fill=diversity_level)) +
  geom_bar(stat = 'identity') +
  labs(title = "Group Constallations and Frequency",
       x = "Groups",
       y = "Frequency",
       fill = "Diversity Level") +
  scale_fill_brewer(palette = "Paired")+
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 8))

p2 #All names are readable on the saved picture
ggsave("p2.png", width = 20, height = 10, bg = 'white')
```

# ANALYSIS

## Diversity Level Performance

### Plotting scores vs diversity
```{r}
df$diversity_level <- as.factor(df$diversity_level)
#synergy
ggplot(data = df, aes(x = diversity_level, y = synergy_score, fill = diversity_level)) + 
  labs(title = "Synergy and Diversity Level")+
  geom_violin()

#team score
team_diversity_plot <- ggplot(data = df, aes(x = diversity_level, y = team_score, fill = diversity_level)) + 
  geom_violin() + 
  geom_boxplot(width = 0.1, outlier.size=-1) + 
  scale_y_continuous(labels = scales::percent_format(accuracy=1))+
  stat_summary(fun = mean, geom = "point", shape = 18, size=2, colour="white", alpha = 0.8)+ 
  guides(fill=FALSE)+
  #scale_fill_manual(name = "diversity level", palette = "Paired")+
  labs(title = "Team Score across Diversity Level", y = "Average Team Score", x = "Diversity")+
  #scale_x_discrete(labels=  c("Low", "Medium", "High")) + 
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), plot.title = element_text(hjust = 0.5, size = 16))


#gain score
gain__diversity_plot <- ggplot(data = df, aes(x = diversity_level, y = gain_score, fill = diversity_level)) + 
  geom_violin() + 
  geom_boxplot(width = 0.1, outlier.size=-1) + 
  scale_y_continuous(labels = scales::percent_format(accuracy=1))+
  stat_summary(fun = mean, geom = "point", shape = 18, size=2, colour="white", alpha = 0.8)+ 
  guides(fill=FALSE)+
  #scale_fill_manual(name = "diversity level", palette = "Paired")+
  labs(title = "Gain Score across Diversity Level", y = "Average Gain Score", x = "Diversity")+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), plot.title = element_text(hjust = 0.5, size = 16))

team_diversity_plot
gain__diversity_plot
```

### Diversity Level and Number of Rounds
```{r}
mean_data <- df %>% 
  group_by(diversity_level) %>% 
  summarise(mean_avg_n_steps = (mean(n_steps, na.rm = TRUE)), sd_mean_avg_n_steps = (sd(n_steps, na.rm = TRUE)), error_mean_avg_n_steps = (std.error(n_steps, na.rm = TRUE)))

tbl1 <- tableGrob(mean_data, theme=ttheme_minimal(), rows=NULL)

wrap_plots(tbl1)

df <- df%>%
  mutate(diversity_level=as.integer(diversity_level))

ggplot(df, aes(diversity_level, n_steps))+
  geom_point()+
  geom_smooth(method='lm', formula= y~x)
```

## Different Group Constallations
```{r}
# Create the grouped bar plot using ggplot2 without labels
p1 <- ggplot(freq_table, aes(x = reorder(lists_str, -gain_score), y = gain_score, fill=diversity_level)) +
  geom_bar(stat = 'identity') +
  labs(title = "Group Constallations and Gain Scores",
       x = "Groups",
       y = "Gain_score",
       fill = "Diversity Level") +
  scale_fill_brewer(palette = "Paired")+
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 8))

p1 #All names are readable on the saved picture
ggsave("p1.png", width = 20, height = 10, bg = 'white')
```

## Other

### Gain score and Number of Discussion Rounds
```{r}
#accuracy and number of ticks/rounds
ggplot(df, aes(x = gain_score, y = n_steps)) + 
  labs(title = "Gain score and Number of Discussion Rounds")+
  geom_point(aes(color=diversity_level))+
  geom_smooth(method='lm', formula= y~x) + 
  theme_bw()
```
## Personality Traits and Group Performance

###Cheking for normality
```{r}
round(pastecs::stat.desc(cbind(team_score=df$team_score[1:4000], synergy_score=df$synergy_score[1:4000], gain_score=df$gain_score[1:4000], neuro=df$neuro_dom[1:4000], know=df$know_dom[1:4000], agree=df$agree_dom[1:4000], crit=df$crit_dom[1:4000], talk=df$talk_dom[1:4000]), basic = FALSE, norm = TRUE), digits = 2)
```

### Correlation Plot
```{r}
df$diversity_level <- as.integer(df$diversity_level)
df_cor <- df %>%
  select(n_steps, team_score, gain_score, synergy_score, diversity_level, agree_dom_0, crit_dom_0, talk_dom_0, neuro_dom_0, know_dom_0)%>%
  cor()
  
df_cor%>%
  corrplot(method="number", number.cex=0.6)

df_cor%>%
  corrplot(method="circle", number.cex=0.6)

df_cor <- round(df_cor,2)

data.frame(df_cor)
```


```{r}
m1 <- stan_glm(gain_score ~ agree_dom_std+talk_dom_std+crit_dom_std+ neuro_dom_std+ know_dom_std, data=df, refresh = 0)
summary(m1, digits = 5)
```

The summary shows that for every 1 step increase of the independent variable, the 
- agree-predictor contributes with -0.02  to the slope
- talk-predictor contributes with  0.1 to the slope 
etc.

```{r }
b1 <- ggplot(df, aes(agree_dom_std, gain_score)) +
  geom_point() +
  geom_abline(
    intercept = coef(b_pred2)[1],
    slope = sum(coef(b_pred2)[2]),
    color = "red") +
  labs(x = "Agree", y = "Gain_score")

b2 <- ggplot(df, aes(talk_dom_std, gain_score)) +
  geom_point() +
  geom_abline(
    intercept = coef(b_pred2)[1],
    slope = sum(coef(b_pred2)[3]),
    color = "red") +
  labs(x = "Talk", y = "Gain_score")

b3 <- ggplot(df, aes(crit_dom_std, gain_score)) +
  geom_point() +
  geom_abline(
    intercept = coef(b_pred2)[1],
    slope = sum(coef(b_pred2)[4]),
    color = "red") +
  labs(x = "Critical", y = "Gain_score")

b4 <- ggplot(df, aes(neuro_dom_std, gain_score)) +
  geom_point() +
  geom_abline(
    intercept = coef(b_pred2)[1],
    slope = sum(coef(b_pred2)[5]),
    color = "red") +
  labs(x = "Neuroticism", y = "Gain_score")

b5 <- ggplot(df, aes(know_dom_std, gain_score)) +
  geom_point() +
  geom_abline(
    intercept = coef(b_pred2)[1],
    slope = sum(coef(b_pred2)[6]),
    color = "red") +
  labs(x = "Knowledge", y = "Gain_score")


plot_grid(b1, b2, b3, b4, b5, ncol=2)
```
```{r}
# Residual versus fitted values plot
plot(fitted(m1),resid(m1))
abline(0,0)
```
From this plot we see that the residual vs fitted values are equally distributed.


### A task for inspecting specific situations
```{r}
### a) Instructor A is a 50-year-old woman who is a native English speaker and has a beauty score of -1. Instructor B is a 60-year-old man who is a native English speaker and has a beauty score of -0.5. Simulate 1000 random draws of the course evaluation rating of these two instructors. In your simulation, use posterior_predict to account for the uncertainty in the regression parameters as well as predictive uncertainty. (Lelia & Anne)

m2 <- stan_glm(gain_score ~ agree_dom_0+talk_dom_0+crit_dom_0+ neuro_dom_0+ know_dom_0, data=df, refresh = 0)
summary(m2, digits = 5)

# Creating a data frame with the informations regarding the two people
df_1 <- df%>%
  select(gain_score, agree_dom_0, talk_dom_0,crit_dom_0,neuro_dom_0, know_dom_0)%>%
  filter(agree_dom_0>2, crit_dom_0>0.1)

# Creating another data frame where the information is used to simulate a 
# posterior prediction with 1000 draws of the evaluation score for each of the 
# instructors
pred1 <- as.data.frame(posterior_predict(m2, newdata=df_1))

pred1 <- dplyr::filter(pred1, row_number()<1001)

head(pred1)

### b) Make a histogram of the difference between the course evaluations for A and B. What is the probability that A will have a higher evaluation? (Mads)

#Calculating the difference between the course evaluations for A and B
pred1$difference <- pred1$'1'-pred1$'2'

# Creating a histogram of the differences between the course evaluations for A 
#and B
ggplot(pred1, aes(difference))+
geom_histogram(aes(y=..density..), colour = "black", fill = "aliceblue", binwidth=  0.05)+
  theme_minimal()+
  stat_function(fun=dnorm,args=list(mean=mean(pred1$difference),sd = sd(pred1$difference)),colour = "red", size =1)+
  geom_vline(xintercept = 0, color="Blue", size = 1) +
  ggtitle("Histogram for difference") 


# Creating a function filtering out all the times A's evaluation is higher than 
#B's
a <- length(Filter(function(diff) { 
    diff > 0 }, 
    pred1$difference))  # Calculating the percentage the weight exeeds the limit 
    
# Calculating the percentage
a/1000

There's a ~45% chance that A will have a higher evaluation scores than B.
```