---
title: "ABM Group Dynamics in CPS"
author: "Kristiane Uhrenholt Warncke and Anne Skamris Holm"
date: "2024-05-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
pacman::p_load(tidyverse, ggpubr, ggalt, ggExtra,  plotrix, dplyr, ggplot2, stringr, tidyr, gridExtra, grid, purrr, lme4)
```

## Loading the data 
```{r}
df <- read.csv2("~/Cognitive_Science/fourth_sem/SocialCulturalDynamics/EXAM/Soccult_ABM/abm-cps-master/ABM_FINAL_trial.csv")
```

### PREPROCESSING

```{r}
#Renaming columns 
df <- rename(df, 
                     round = X.run.number., 
                     team_size = team_size, 
                     game_size = game_size,
                     know_sharing = knowledge_sharing,
                     n_steps = X.step.,
                     team_score = team_score,
                     gain_score = gain_score,
                     synergy_score = synergy_score,
                     agent_names = agent_names,
                     reached_consensus = reached_consensus.,
                     agents_agree = X.agreeableness..of.turtles,
                     agents_talk = X.talkativeness..of.turtles,
                     agents_neuro = X.neuroticism..of.turtles,
                     agents_crit = X.critical_thinking..of.turtles,
                     agents_know= X.knowledge..of.turtles)
```

## Converting string vectors to numeric vectors in certain columns
```{r}
# Helper function to convert string vectors to numeric vectors
convert_to_numeric_vector <- function(x) {
  # Remove unwanted characters
  cleaned <- gsub("\\[|\\]|\"", "", x)
  # Split the string by spaces and convert to numeric
  as.numeric(unlist(strsplit(cleaned, " ")))
}

# Apply this conversion to each agents_ column
df <- df %>%
  mutate(across(starts_with("agents_"), ~map(.x, convert_to_numeric_vector)))

# ..and to the synergy score
df <- df %>%
  mutate(across("synergy_score", ~map(.x, convert_to_numeric_vector)))

# and team score
df <- df %>%
  mutate(across("team_score", ~map(.x, convert_to_numeric_vector)))

# and gain score
df <- df %>%
  mutate(across("gain_score", ~map(.x, convert_to_numeric_vector)))

```
## Grouping the diversity_score into levels from 1-5 depending on team size 

```{r}
# Calculate diversity level
df <- df %>%
  mutate(
    agent_names_cleaned = str_extract_all(agent_names, "\\b\\w+_agent\\b"), # Extract agent types
    diversity_level = map_int(agent_names_cleaned, ~length(unique(.x))) # Count unique agent types
  )%>%
  select(-know_sharing, -team_size, -game_size, -agent_names)
```

```{r}
df <- df %>%
  mutate(
    round = as.integer(round),
    n_steps = as.numeric(n_steps),
    team_score = as.numeric(team_score),
    gain_score = as.numeric(gain_score),
    synergy_score = as.numeric(synergy_score),
    agent_names_cleaned = as.character(agent_names_cleaned),
    reached_consensus = as.logical(reached_consensus),  # this is a TRUE/FALSE column
    diversity_level = as.factor(diversity_level)
  )

df <- df %>%
  drop_na() %>%
  mutate(agree_dom = as.numeric(lapply(agents_agree, mean)))%>%
  mutate(talk_dom = as.numeric(lapply(agents_talk, mean)))%>%
  mutate(crit_dom = as.numeric(lapply(agents_crit, mean)))%>%
  mutate(know_dom = as.numeric(lapply(agents_know, mean)))%>%
  mutate(neuro_dom = as.numeric(lapply(agents_neuro, mean)))

df <- df %>%
  mutate(agree_dom= ifelse(agree_dom==0, NA, agree_dom))%>%
  mutate(talk_dom = ifelse(talk_dom==0, NA, talk_dom))%>%
  mutate(crit_dom = ifelse(crit_dom==0, NA, crit_dom))%>%
  mutate(know_dom = ifelse(know_dom==0, NA, know_dom))%>%
  mutate(neuro_dom = ifelse(neuro_dom==0, NA, neuro_dom))

df <- df %>%
  mutate(agree_dom_std = scale(df$agree_dom))%>%
  mutate(talk_dom_std = scale(talk_dom))%>%
  mutate(crit_dom_std = scale(crit_dom))%>%
  mutate(know_dom_std = scale(know_dom))%>%
  mutate(neuro_dom_std = scale(neuro_dom))

df <- df %>%
  select(round, n_steps, team_score, gain_score, synergy_score, reached_consensus, diversity_level, agents_agree,agree_dom, agree_dom_std,  agents_talk,talk_dom,talk_dom_std,  agents_crit,crit_dom,crit_dom_std,agents_know,know_dom,know_dom_std,agents_neuro, neuro_dom,neuro_dom_std, agent_names_cleaned
         )%>%
  filter(reached_consensus=="TRUE")
```


# Data visualization

```{r}
#Team, gain and synergy plots
synergyscore_plot <- ggplot(df, aes(x = synergy_score)) +
  geom_histogram(bins = 30, fill = "skyblue", color = "black") +
  theme_minimal() +
  labs(title = "Distribution of Group Performance on the three score-types", x = "Synergy Score", y = "Frequency")

teamscore_plot <- ggplot(df, aes(x = team_score)) +
  geom_histogram(bins = 30, fill = "lightgreen", color = "black") +
  theme_minimal() +
  labs(title = "", x = "Team Score", y = "")

gainscore_plot <- ggplot(df, aes(x = gain_score)) +
  geom_histogram(bins = 30, fill = "lightpink", color = "black") +
  theme_minimal() +
  labs(title = "", x = "Gain score", y = "")

grid.arrange(synergyscore_plot, teamscore_plot, gainscore_plot, ncol=3)
```

```{r}
df%>% 
  group_by(diversity_level) %>% 
  summarise("count" = n())

#Diversity level plot
ggplot(data = df, aes(x = diversity_level, fill = diversity_level)) + 
  geom_histogram(stat = "count") + 
  labs(title = "Distribution of Diversity Levels", x = "Diversity Level", y = "Count") +
  scale_fill_brewer(palette = "Paired") +
  theme_bw()
```


```{r}
#distribution of agents
colSums(!is.na(df))[8:22]
```

### density plots....

```{r}
#long data to make density plots across traits
long_data <- df %>%
  pivot_longer(
    cols = c("talk_dom_std", "neuro_dom_std", "crit_dom_std", "know_dom_std", "agree_dom_std"),
    names_to = "trait",
    values_to = "value"
  ) %>%
  drop_na(value)

density_plot_combined <- ggplot(long_data, aes(x = value, fill = trait, color = trait)) +
  geom_density(alpha = 0.7) +  # Adding transparency to see overlapping areas
  scale_fill_brewer(palette = "Paired") +  # This sets distinct colors for each trait
  scale_color_brewer(palette = "Paired") +
  theme_minimal() +
  labs(
    title = "Distribution of Scaled Dominating Traits",
    x = "Trait Value",
    y = "Density"
  ) +
  scale_x_continuous(limits = c(-4, 6)) +
  theme(legend.title = element_blank())  # Optionally remove the legend title

density_plot_combined
```

```{r}
ggplot(long_data, aes(x = trait, y = value, fill = trait)) +
  geom_boxplot() +
  scale_fill_brewer(palette = "Set2") +  # Using a color palette for distinction
  theme_minimal() +
  labs(title = "Boxplots of Trait Values", x = "Trait", y = "Value") +
  scale_y_continuous(limits = c(-2, 7)) +
  theme(legend.position = "none")  # Optional: remove the legend if not needed
```

```{r}
#Calculating average synergy, team and gain scores
average_synergy_score <- mean(df$synergy_score)
average_team_score <- mean(df$team_score)
average_gain_score <- mean(df$gain_score)

# Print the result
print(paste("Average synergy score:", average_synergy_score))
print(paste("Average team score:", average_team_score))
print(paste("Average gain score:", average_gain_score))

```

```{r}
mean_agree <- mean(df$agree_dom, na.rm = TRUE)
mean_talk <- mean(df$talk_dom, na.rm = TRUE)
mean_neuro <- mean(df$neuro_dom, na.rm = TRUE)
mean_crit <- mean(df$crit_dom, na.rm = TRUE)
mean_know <- mean(df$know_dom, na.rm = TRUE)

# Print the results
print(paste("Average agree:", mean_agree))
print(paste("Average talk:", mean_talk))
print(paste("Average neuro:", mean_neuro))
print(paste("Average crit:", mean_crit))
print(paste("Average know:", mean_know))
```
# ANALYSIS

```{r}
df <- df %>%
  mutate(diversity_level = as_factor(diversity_level))
```


##Plotting scores vs diversity
```{r}
#synergy
ggplot(data = df, aes(x = diversity_level, y = synergy_score, fill = diversity_level)) + 
  labs(title = "Synergy and Diversity Level")+
  geom_violin()


#team score
team_diversity_plot <- ggplot(data = df, aes(x = diversity_level, y = team_score, fill = diversity_level)) + 
  geom_violin() + 
  geom_boxplot(width = 0.1, outlier.size=-1) + 
  #scale_y_continuous(labels = scales::percent_format(accuracy=1))+
  stat_summary(fun = mean, geom = "point", shape = 18, size=2, colour="white", alpha = 0.8)+ 
  guides(fill=FALSE)+
  scale_fill_brewer(palette = "Paired")+
  labs(title = "Team Score across Diversity Level", y = "Average Team Score", x = "Diversity")+
  #scale_x_discrete(labels=  c("Low", "Medium", "High")) + 
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), plot.title = element_text(hjust = 0.5, size = 16))


#gain score
gain__diversity_plot <- ggplot(data = df, aes(x = diversity_level, y = gain_score, fill = diversity_level)) + 
  geom_violin() + 
  geom_boxplot(width = 0.1, outlier.size=-1) + 
  #scale_y_continuous(labels = scales::percent_format(accuracy=1))+
  stat_summary(fun = mean, geom = "point", shape = 18, size=2, colour="white", alpha = 0.8)+ 
  guides(fill=FALSE)+
  scale_fill_brewer(palette = "Paired")+
  labs(title = "Gain Score across Diversity Level", y = "Average Team Score", x = "Diversity")+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), plot.title = element_text(hjust = 0.5, size = 16))

team_diversity_plot
gain__diversity_plot
```




```{r}


mean_data <- df %>% 
  group_by(diversity_level) %>% 
  summarise(mean_avg_n_steps = (mean(n_steps, na.rm = TRUE)), sd_mean_avg_n_steps = (sd(n_steps, na.rm = TRUE)), error_mean_avg_n_steps = (std.error(n_steps, na.rm = TRUE)))

mean_data
```


```{r}
df <- df%>%
  mutate(diversity_level=as.integer(diversity_level))

ggplot(df, aes(diversity_level, n_steps))+
  geom_point()+
  geom_smooth(method='lm', formula= y~x)
```


```{r}
#accuracy and number of ticks/rounds

ggplot(df, aes(x = team_score, y = n_steps, color = diversity_level)) + 
  labs(title = "Team score and Number of Discussion Rounds", y = "Number of Discussion Rounds", x = "Team Score")+
  geom_point()+
  geom_smooth(method='lm', formula= y~x) + 
  scale_fill_brewer(palette = "Paired") +
  theme_bw()
```



```{r}
#tics and diversity
ticks_diversity_plot <- ggplot(data = mean_data, aes(x = diversity_level, y = mean_avg_n_steps, color = diversity_level, ymin = (mean_avg_n_steps - 2*error_mean_avg_n_steps), ymax = (mean_avg_n_steps + 2*error_mean_avg_n_steps))) +
  geom_errorbar(width = 0.5, size = 1.5)+
  ylim(16, 19.5) + 
  geom_point(size = 3, shape = 18, color = "black") +
  labs(title = "Discussion Rounds", y = "Average Number of Discussion Rounds", x = "Diversity")+
    coord_equal(ratio = 2)+
  guides(color=FALSE)+
  theme_bw()
  #theme(plot.title = element_text(hjust = 0.5, size = 16), text=element_text(family = "Times New Roman"))

ticks_diversity_plot
```


##Correlation plot

##H2.7: Gain scores and synergy scores are correlated

Annes com

## STATS 

### H.1.1 Groups with high diversity levels will on average have higher gain scores
```{r}
#take out diversity level 1, as it has too few groups (6 groups) to base anything on
df <- df %>%
  filter(diversity_level != 1)

# team score/accuracy model
team_score_model <- lm(team_score ~ diversity_level, data = df)
summary(team_score_model)

# synergy model
synergy_score_model <- lm(synergy_score ~ diversity_level, data = df)
summary(synergy_score_model)

# gain score model
gain_score_model <- lm(gain_score ~ diversity_level, data = df)
summary(gain_score_model)
```




## H2.3: Groups with more agents with the dominating trait agreeableness will on average decrease number of rounds

## H2.4: Groups with more agents with the dominating trait talkativeness will on average increase number of rounds
```{r}
rounds_trait_model <- lm(n_steps ~ agree_dom + talk_dom + crit_dom + know_dom + neuro_dom, data = df)
summary(rounds_trait_model)
```

##H2.2: Groups with more agents with the dominating traits agreeableness and talkativeness will on average increase gain- and synergy scores

##H2.5: Groups with more agents with the dominating trait critical thinking will on average increase synergy scores

##H2.6: Critical thinking as a personality trait has low influence on gain scores

##H2.9: Groups with more agents with the dominating trait neuroticism will on average decrease synergy and gain scores

```{r}
gain_trait_model <- lm(gain_score ~ agree_dom + talk_dom + crit_dom + know_dom + neuro_dom, data = df)
summary(gain_trait_model)

synergy_trait_model <- lm(synergy_score ~ agree_dom + talk_dom + crit_dom + know_dom + neuro_dom, data = df)
summary(synergy_trait_model)
```
```


